\textbf{\supplement}\vskip 2ex \mytitle\vskip 2ex

### 2023-05-12 generated/over-fitting-code-section.tex
Mainly getting mathematica+make to work together
Getting make and mathematica to run together properly
Tidied up programs and scripts
Revisions for Computer Journal

\noindent
Poor code undermines scientific quality: poor code can generate plausible results from \emph{any\/} data, including even fictional science. A temptation is that %artificial world \cite{sciences-artificial}, built on
adjusting the code's results gets more attention than the theoretical basis of the code's overall faithfulness to the experimental phenomena. 
\def\imageWidthCalculation{\multiply \imageWidth by 21 \divide \imageWidth by 28}
\def\figureStarDetails{*}

\begin{figure\figureStarDetails}[t]
{ \newdimen \imageWidth 
  \imageWidth=\textwidth
  \imageWidthCalculation
  \begin{center}\includegraphics[width=\imageWidth]{
generated/mathematicaplot.pdf
}\end{center}}
\caption{Much computational science is concerned with finding plausible multi-dimensional models that fit models to data with the aim of extrapolating or predicting new results from them. Shown here is notional sample of experimental 2D data (the dots), a linear least squares regression, and an exact polynomial model. The over-fitted polynomial model fits the sample \emph{exactly\/}, but since the experimental data is presumably subject to random error (indicated by the confidence interval, itself estimated) the linear model would generally be considered a better description of the experimental data.}
\label{fig-overfit}
\end{figure\figureStarDetails}

In modeling terms, superficially successful computer programs tend to \emph{over-fit\/} phenomena \cite{over-fit}. Instead of using code to test the models, we tinker with the code adapting it closer and closer to our prejudices. The code then apparently confirms our science, since we fitted it to our understanding.

@@

Unfortunately, code in published science \emph{is\/} often over-fitted, and over-fitted in a way that is very hard to scrutinize. For example, in epidemiology (which is considered in section \ref{section-pandemic-modeling}) it is routine to use very complicated, large dynamical models parameterized with numerous social, cultural, health, demographic and geographical data. The parameterization is mixed between data files, data written explicitly into the code, and with conditionals and other structuring in the code to cover special cases. Indeed, many of the programs in the survey used comment to inactivate code, presumably indicating an unfinished tinkering approach to code development.\footnote{Of the \the\countRepos\ papers in the pilot survey that reported use of code repositories (covering \input generated/kloc.tex
\totalkLOC\ thousand lines of code --- so this is not a trivial amount of programming effort), \numbertoname{\the\countMissingCode} provided an empty repository with no code at all (effectively commenting out all their code!), and \numbertoname{\the\countCommentedOut} repositories explicitly commented out chunks of workable code. The \numbertoname{\the\countNotCommentedOut} remaining non-trivial repositories with no commented-out code consisted of straightforward, short code files with few comments of any sort.\remembernamevalue{numbernamesexamplefootnote}{\thefootnote}}

@@

\vfill % the vfill is only needed to put the box at the bottom of a column if that makes it look better
\vskip 1ex
\noindent 
{\newdimen\boxwidth
\boxwidth = \columnwidth
\advance\boxwidth by - 1em
\begin{tabular}{@{}l@{}}
\framebox[\columnwidth][r]{\hfill\parbox{\boxwidth}{This section (\ref{over-fit}) of this paper, including figure \ref{fig-overfit}, was generated as a computational paper in a \emph{Mathematica\/} notebook, then exported to \LaTeX\ to include here. The \supplement\ provides a more complete explanation, as well as the data and details of the fitted polynomials (also generated automatically). This paper's repository includes the notebook.

>>>>

\noindent
Poor code undermines scientific quality: poor code can generate plausible and possibly misleading results from \emph{any\/} data, including even fictional science. A temptation is that %artificial world \cite{sciences-artificial}, built on
adjusting the code's results gets more attention than the theoretical basis of the code's overall faithfulness to the experimental phenomena. 
% wolframscript (if being used) unnecessarily prints the next 2 lines (generated by CellPrint), but they are needed!
\def\imageWidthCalculation{\multiply \imageWidth by 21 \divide \imageWidth by 28}
\def\figureStarDetails{*}

% wolframscript (if being used) unnecessarily prints the next 2 lines (generated by CellPrint), but they are needed because the text also ends up in notebook cells!
\def\imageWidthCalculation{\multiply \imageWidth by 21 \divide \imageWidth by 28}
\def\figureStarDetails{*}

\begin{figure\figureStarDetails}[t]
{ \newdimen \imageWidth 
  \imageWidth=\textwidth
  \imageWidthCalculation
  \begin{center}\includegraphics[width=\imageWidth]{
generated/mathematicaplot.jpg
}\end{center}}
\caption{Much computational science is concerned with finding plausible multi-dimensional models that fit models to data with the aim of extrapolating or predicting new results from them. Shown here is notional sample of experimental 2D data (the dots), a linear least squares regression, and an exact polynomial model. The over-fitted polynomial model fits the sample \emph{exactly\/}, but since the experimental data is presumably subject to random error (indicated by the confidence interval, itself estimated) the linear model would generally be considered a better description of the experimental data.}
\label{fig-overfit}
\end{figure\figureStarDetails}

In modeling terms, superficially successful computer programs tend to \emph{over-fit\/} phenomena \cite{over-fit}. Instead of using code to test the models, we tinker with the code adapting it to be closer and closer to our prejudices. The code then apparently confirms our science, since we fitted it to our understanding.

@@

Unfortunately, code in published science \emph{is\/} often over-fitted, and over-fitted in a way that is very hard to scrutinize. For example, in epidemiology (which is considered in section \ref{section-pandemic-modeling}) it is routine to use very complicated, large dynamical models parameterized with numerous social, cultural, health, demographic and geographical data. The parameterization is mixed between data files, data written explicitly into the code, and with conditionals and other structuring in the code to cover special cases. Indeed, many of the programs in the survey used comments to inactivate code, presumably indicating an unfinished tinkering approach to code development.\footnote{Of the \the\countUsesVersionControlRepository\ papers in the pilot survey that reported use of code repositories (covering \input generated/kloc.tex
\totalkLOC\ thousand lines of code --- so this is not a trivial amount of programming effort), \numbertoname{\the\countMissingCode} provided an empty repository with no code at all (effectively commenting out all their code!), and \numbertoname{\the\countCommentedOut} repositories explicitly commented out chunks of workable code. The \numbertoname{\the\countNotCommentedOut} remaining non-trivial repositories with no commented-out code consisted of straightforward, short code files with few comments of any sort.\remembernamevalue{numbernamesexamplefootnote}{\thefootnote}}

@@

%\vfill % the vfill is only needed to put the box at the bottom of a column if that makes it look better
\vskip 1ex
\noindent 
{\newdimen\boxwidth
\boxwidth = \columnwidth
\advance\boxwidth by - 1em
\begin{tabular}{@{}l@{}}
\framebox[\columnwidth][r]{\hfill\parbox{\boxwidth}{This section (\ref{over-fit}) of this paper, including figure \ref{fig-overfit}, was generated as a computational paper in a \emph{Mathematica\/} notebook, then exported to \LaTeX\ to include here. The \supplement\ provides a more complete explanation, as well as the data and details of the fitted polynomials (also generated automatically). This paper's repository includes the notebook.

### 2023-05-12 generated/repos.tex
Various corrections inspired by Konrad Hinsen
Mainly getting mathematica+make to work together
Revisions for Computer Journal

\begin{tabular}{|l|r|rrr@{\hskip .75ex}l|}\hline
&\multicolumn{1}{c|}{\bf PDF paper}&\multicolumn{4}{c|}{\bf Repository code \& data}\\ \cline{2-6}
\multicolumn{1}{|c|}{\bf Github repository}&\multicolumn{1}{c|}{\bf Number}&\multicolumn{1}{c}{\bf Number}&\multicolumn{1}{c}{\bf Code}&\multicolumn{2}{c|}{\bf Data}\\
\multicolumn{1}{|c|}{\bf and paper citation}&\multicolumn{1}{c|}{\bf of pages}&\multicolumn{1}{c}{\bf of files}&\multicolumn{1}{c}{\bf kLOC}&\multicolumn{2}{c|}{\bf bytes} \\ \hline\hline
\reponame{AI-CDSS-Cardiovascular-Silo}{AI-CDSS-Cardiovascular-Silo} & \csname pagelength-AI-CDSS-Cardiovascular-Silo\endcsname & \commarise{206} & 143 & 64&Mb \\
\reponame{blast-ct}{blast-ct} & \csname pagelength-blast-ct\endcsname & \commarise{44} & 2 & 87&Mb \\
\reponame{covid-sim}{covid-sim} & \csname pagelength-covid-sim\endcsname & \commarise{229} & 25 & 734&Mb \\
\reponame{lactModel}{lactModel} & \csname pagelength-lactModel\endcsname & \commarise{20} & 2 & 165&kb \\
\reponame{LRM}{LRM} & \csname pagelength-LRM\endcsname & \commarise{125} & 8 & 2&Mb \\
\reponame{manifold-ga}{manifold-ga} & \csname pagelength-manifold-ga\endcsname & \commarise{11} & 1 & \multicolumn{2}{c|}{---} \\
\reponame{MetricSelectionFramework}{MetricSelectionFramework} & \csname pagelength-MetricSelectionFramework\endcsname & \commarise{44} & 4 & 236&kb \\
\reponame{PENet}{PENet} & \csname pagelength-PENet\endcsname & \commarise{115} & 8 & 3&Mb \\
\reponame{philter-ucsf}{philter-ucsf} & \csname pagelength-philter-ucsf\endcsname & \commarise{1987} & 13 & 32&Mb \\
\reponame{PostoperativeOutcomes\_RiskNet}{PostoperativeOutcomes.RiskNet} & \csname pagelength-PostoperativeOutcomes.RiskNet\endcsname & \commarise{1} & --- & \multicolumn{2}{c|}{---} \\
\reponame{SiameseChange}{SiameseChange} & \csname pagelength-SiameseChange\endcsname & \commarise{5} & 1 & 1&kb \\
\hline \multicolumn{1}{|r|}{{\bf Average} ($N=11$)}&\makeAverage{\the\gitPages}{11}&253& 19& 84&Mb\\
\hline \end{tabular}

>>>>

\begin{tabular}{|l|r|rrr@{\hskip .75ex}l|}\hline
&\multicolumn{1}{c|}{\bf PDF paper}&\multicolumn{4}{c|}{\bf Repository code \& data}\\ \cline{2-6}
\multicolumn{1}{|c|}{\bf Github repository}&\multicolumn{1}{c|}{\bf Number}&\multicolumn{1}{c}{\bf Number}&\multicolumn{1}{c}{\bf Code}&\multicolumn{2}{c|}{\bf Data}\\
\multicolumn{1}{|c|}{\bf and paper citation}&\multicolumn{1}{c|}{\bf of pages}&\multicolumn{1}{c}{\bf of files}&\multicolumn{1}{c}{\bf kLOC}&\multicolumn{2}{c|}{\bf bytes} \\ \hline\hline
\reponame{AI-CDSS-Cardiovascular-Silo}{AI-CDSS-Cardiovascular-Silo} & \csname pagelength-AI-CDSS-Cardiovascular-Silo\endcsname & \digitBlock{206} & 143 & 64&Mb \\
\reponame{blast-ct}{blast-ct} & \csname pagelength-blast-ct\endcsname & \digitBlock{54} & 3 & 238&Mb \\
\reponame{covid-sim}{covid-sim} & \csname pagelength-covid-sim\endcsname & \digitBlock{229} & 25 & 734&Mb \\
\reponame{lactModel}{lactModel} & \csname pagelength-lactModel\endcsname & \digitBlock{20} & 2 & 165&kb \\
\reponame{LRM}{LRM} & \csname pagelength-LRM\endcsname & \digitBlock{125} & 8 & 2&Mb \\
\reponame{manifold-ga}{manifold-ga} & \csname pagelength-manifold-ga\endcsname & \digitBlock{11} & 1 & \multicolumn{2}{c|}{---} \\
\reponame{MetricSelectionFramework}{MetricSelectionFramework} & \csname pagelength-MetricSelectionFramework\endcsname & \digitBlock{44} & 4 & 236&kb \\
\reponame{PENet}{PENet} & \csname pagelength-PENet\endcsname & \digitBlock{117} & 8 & 4&Mb \\
\reponame{philter-ucsf}{philter-ucsf} & \csname pagelength-philter-ucsf\endcsname & \digitBlock{1987} & 13 & 32&Mb \\
\reponame{PostoperativeOutcomes\_RiskNet}{PostoperativeOutcomes.RiskNet} & \csname pagelength-PostoperativeOutcomes.RiskNet\endcsname & \digitBlock{1} & --- & \multicolumn{2}{c|}{---} \\
\reponame{SiameseChange}{SiameseChange} & \csname pagelength-SiameseChange\endcsname & \digitBlock{5} & 1 & 1&kb \\
% Numbers below are rounded, but more precise averages are 254.454545 -- 18824.000000 -- 97771458.363636 = 1075486042 / 11
\hline \multicolumn{1}{|r|}{{\bf Average} ($N=11$)}&\makeAverage{\the\gitPages}{11}&254& 19& 98&Mb\\
\hline \end{tabular}

### 2023-05-12 generated/summary-table.tex
Revisions for Computer Journal

\begin{tabular}{|rrrc|}\hline
Number of papers sampled relying on code&32&100\%&\\\hline\hline
\multicolumn{4}{|l|}{\textbf{Access to code}}\\
Some or all code available&12&38\%&\\
Some or all code in principle available on request&8&25\%&\\
Requested code actually made available (within 2 years 8 months${}^{\star}$)&0&0\%& \\\hline
 \multicolumn{4}{|l|}{{\textbf{Evidence of any software engineering practice}}}\\
Evidence program designed rigorously&{0}&{0\%}&\\
Evidence source code properly tested&{0}&{0\%}&\\
Evidence of any tool-based development&{0}&{0\%}&\\
Team or open source based development&{0}&{0\%}&\\
Other methods, e.g., independent coding methods&{1}&{3\%}&\\\hline
\multicolumn{4}{|l|}{\textbf{Documentation and comments}}\\
Substantial code documentation and comments&2&6\%&\\
Comments explain some code intent&3&9\%&\\
Procedural comments (e.g., author, date, copyright)&10&31\%&\\
No usable comments&17&53\%&\\\hline
\multicolumn{4}{|l|}{\textbf{Repository use}}\\
Used code repository (e.g., GitHub)&9&28\%&\\
Used data repository (e.g., Dryad or GitHub)&9&28\%&\\
Empty repository&1&3\%&\\\hline
\multicolumn{4}{|l|}{{\textbf{Evidence of documented processes}}}\\
Evidence of RAP/\RAPstar\ or any other principles in use to support scrutiny&{0}&{0\%}&\\\hline
\multicolumn{4}{|l|}{\textbf{Adherence to journal code policy (if any)}}\\
Papers published in journals with code policies&26&81\%&\\
Clear breaches of journal code policy (if any) & 11&42\%&(\emph{N}~=~26)\\
\hline\end{tabular}\\
\vskip 1ex ${}^{\star}$Time of 2 years 8 months is wait between code request and date of generating this table.

>>>>

\begin{tabular}{|rrrc|}\hline
Number of papers sampled relying on code&32&100\%&\\\hline\hline
\multicolumn{4}{|l|}{\textbf{Access to code}}\\
Some or all code available&12&38\%&\\
Some or all code in principle available on request&8&25\%&\\
Requested code actually made available (within 2 years 10 months${}^{\star}$)&0&0\%& \\\hline
 \multicolumn{4}{|l|}{{\textbf{Evidence of any software engineering practice}}}\\
Evidence program designed rigorously&{0}&{0\%}&\\
Evidence source code properly tested&{0}&{0\%}&\\
Evidence of any tool-based development&{0}&{0\%}&\\
Team or open source based development&{0}&{0\%}&\\
Other methods, e.g., independent coding methods&{1}&{3\%}&\\\hline
\multicolumn{4}{|l|}{\textbf{Documentation and comments}}\\
Substantial code documentation and comments&2&6\%&\\
Comments explain some code intent&3&9\%&\\
Procedural comments (e.g., author, date, copyright)&10&31\%&\\
No usable comments&17&53\%&\\\hline
\multicolumn{4}{|l|}{\textbf{Repository use}}\\
Used code repository (e.g., GitHub)&9&28\%&\\
Used data repository (e.g., Dryad or GitHub)&9&28\%&\\
Empty repository&1&3\%&\\\hline
\multicolumn{4}{|l|}{{\textbf{Evidence of documented processes}}}\\
Evidence of RAP/\RAPstar\ or any other principles in use to support scrutiny&{0}&{0\%}&\\\hline
\multicolumn{4}{|l|}{\textbf{Adherence to journal code policy (if any)}}\\
Papers published in journals with code policies&26&81\%&\\
Clear breaches of journal code policy (if any) & 11&42\%&(\emph{N}~=~26)\\
\hline\end{tabular}\\
\vskip 1ex ${}^{\star}$Time of 2 years 10 months is wait between code request and date of generating this table.
